---
title: "R Notebook"
output: html_notebook
---

Studentized Residuals
```{r}
url = 
'http://www.statsci.org/data/general/hills.txt' 
races.table = read.table(url, header=TRUE, 
sep='\t')
n.index=seq(1,nrow(races.table))
races.table=cbind(races.table,n.index)

lm.model=lm(Time~Distance+Climb,data=races.table)
ggplot(lm.model,aes(x=n.index,y=rstudent(lm.model)))+geom_point(color="orange")+geom_line(y=-3)+geom_line(y=3)+labs(title = "External Studentized Residuals",x="Observation",y="Residuals")

```
Cook's D
```{r}
##Cook's D
D.cut=4/(nrow(races.table)-3-1)
ggplot(lm.model,aes(x=n.index,y=cooks.distance(lm.model)))+geom_point(color="orange")+geom_line(y=D.cut)+labs(title = "Cook's D",x="Observation",y="Cook's Distance")
```
DFFITS
```{r}
df.cut=2*(sqrt((3+1)/nrow(races.table)))

ggplot(lm.model,aes(x=n.index,y=dffits(lm.model))) +geom_point(color="orange") +geom_line(y=df.cut) +geom_line(y=-df.cut)+labs(title = "DFFITS",x="Observation",y="DFFITS")
```
Hat Values
```{r}
hat.cut=2*(3+1)/nrow(races.table)

ggplot(lm.model,aes(x=n.index,y=hatvalues(lm.model)))+geom_point(color="orange")+geom_line(y=hat.cut)+labs(title = "Hat values",x="Observation",y="Hat Values")
```
DFBETA
```{r}
db.cut=2/sqrt(nrow(races.table))

ggplot(lm.model,aes(x=n.index,y=dfbetas(lm.model)[,'Climb']))+geom_point(color="orange")+
  geom_line(y=db.cut)+geom_line(y=-db.cut)+labs(title = "DFBETA for Climb",x="Observation",y="DFBETAS")

ggplot(lm.model,aes(x=n.index,y=dfbetas(lm.model)[,'Distance']))+geom_point(color="orange")+geom_line(y=db.cut)+
  geom_line(y=-db.cut)+labs(title = "DFBETA for Distance",x="Observation",y="DFBETAS")
```
How to Handle Influential Observations

1. Recheck the data to ensure that no transcription or data entry errors 
occurred.

2. If the data is valid, one possible explanation is that 
the model is not adequate.
 A model with higher-order terms, such as polynomials and 
interactions between the variables, might be necessary to fit the 
data well.
 Nonlinear model

3. Determine the robustness of the inference by running the analysis 
both with and without the influential observations.

4. Robust Regression (Covered Later in Program)

5. Weighted Least Squares (WLS)

```{r}
cor(mtcars)
lm.model=lm(mpg~.,data=mtcars)
vif(lm.model)
```

